\vspace{-\baselineskip} Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-2\baselineskip}

### 2017-06-05 paper/ms.tex

Since the covariance matrix may be written as the sum of exponentials, the predicted mean is given by
\begin{eqnarray}
\bar f^*_m &=& K(t^*_m,\bvec{t}) \bvec{z} \\
&=& \sum_n \sum_{j=1}^J e^{-c_j \vert t^*_m-t_n\vert}\left(a_j \cos{(d_j \vert t^*_m-t_n\vert)} +b_j \sin{(d_j \vert t^*_m-t_n\vert)}\right) z_j,
\end{eqnarray}
for $m =1$ to $M$.
Dividing the sum over $n$ into $\{t_1,...,t_{n_0}\} < t^*_m$ and
$\{t_{n0+1},...,t_N\} \ge t^*_m$, gives
\begin{eqnarray}
\bar f^*_m &=& \sum_{j=1}^J\sum_{n=1}^{n_0} e^{-c_j(t^*_m-t_n)} \left[a_j \cos{(d_j(t^*_m-t_n))}+b_j\sin{(d_j(t^*_m-t_n))}\right] z_n \\
&+& \sum_{j=1}^J\sum_{n=n_0+1}^N  e^{-c_j(t_n-t^*_m)} \left[a_j \cos{(d_j(t_n-t^*_m))}+b_j\sin{(d_j(t_n-t^*_m))}\right] z_n.
\end{eqnarray}
We can compute this recursively.  Defining
\begin{eqnarray}
Q^-_{n_0,i} &=& \left(Q^-_{n_0-1,i}+z_n\tilde V_{n_0,i}\right)e^{-c_j(t_{n+1}-t_n)},\\
X^-_{m,n_0,i} &=& e^{-c_j(t_m^*-t_{n+1})}\tilde U_{m,i},\\
Q^+_{n_0,i} &=& \left(Q^+_{n_0+1,i}+z_n\tilde U_{n,i}\right)e^{-c_j(t_n-t_{n-1})},\\
X^+_{m,n,i} &=& e^{-c_j(t_{n-1}-t_m^*)} \tilde V_{m,i},
\end{eqnarray}
where $Q^-_{0,i} = 0$ and $Q^+_{N+1,i}=0$ for $i=1$ to $2J$, and in $X^\pm_{m,n,i}$ the expressions
for $\tilde U_{m,i}$ and $\tilde V_{m,i}$ are evaluted at $t_m^*$.
Two sweeps must be carried out, upward in $n$ and $m$ to compute $Q^-$ recursively, and downward in
$m$ and $n$ to compute $Q^+$ recursively, for a total scaling of $\approx \mathcal{O}(J^2N+J(N+M))$.
For each value of $m$, $Q^\pm$ are recursively updated over $n$ until $n_0$ is reached, at which
point the predicted mean can be computed from
\begin{equation}
\bar f^*_i = \sum_{i=1}^{2J} \left[Q^-_{n_0,i}X^-_{m,n_0,i}+Q^+_{n_0,i}X^+_{m,n_0,i}\right],
\end{equation}

>>>>

Since the covariance matrix may be written as the sum of exponentials, the predicted mean is given by
\begin{eqnarray}
\bar f^*_m &=& K(t^*_m,\bvec{t}) \bvec{z} \\
&=& \sum_n \sum_{j=1}^J e^{-c_j \vert t^*_m-t_n\vert}\left(a_j \cos{(d_j \vert t^*_m-t_n\vert)} +b_j \sin{(d_j \vert t^*_m-t_n\vert)}\right) z_j,
\end{eqnarray}
for $m =1$ to $M$.
Dividing the sum over $n$ into $\{t_1,...,t_{n_0}\} < t^*_m$ and
$\{t_{n0+1},...,t_N\} \ge t^*_m$, gives
\begin{eqnarray}
\bar f^*_m &=& \sum_{j=1}^J\sum_{n=1}^{n_0} e^{-c_j(t^*_m-t_n)} \left[a_j \cos{(d_j(t^*_m-t_n))}+b_j\sin{(d_j(t^*_m-t_n))}\right] z_n \\
&+& \sum_{j=1}^J\sum_{n=n_0+1}^N  e^{-c_j(t_n-t^*_m)} \left[a_j \cos{(d_j(t_n-t^*_m))}+b_j\sin{(d_j(t_n-t^*_m))}\right] z_n.
\end{eqnarray}
We can compute this recursively.  Defining
\begin{eqnarray}
Q^-_{n,i} &=& \left(Q^-_{n-1,i}+z_n\tilde V_{n,i}\right)e^{-c_j(t_{n+1}-t_{n})},\\
X^-_{m,n,i} &=& e^{-c_j(t_m^*-t_{n+1})}\tilde U_{m,i},\\
Q^+_{n,i} &=& \left(Q^+_{n+1,i}+z_n\tilde U_{n,i}\right)e^{-c_j(t_{n}-t_{n-1})},\\
X^+_{m,n,i} &=& e^{-c_j(t_{n-1}-t_m^*)} \tilde V_{m,i},
\end{eqnarray}
where $t_0=t_1$, $t_{N+1}=t_N$, $Q^-_{0,i} = 0$ and $Q^+_{N+1,i}=0$ for $i=1$ to $2J$, and in $X^\pm_{m,n,i}$ the expressions
for $\tilde U_{m,i}$ and $\tilde V_{m,i}$ are evaluted at $t_m^*$.
Two sweeps must be carried out, upward in $n$ and $m$ to compute $Q^-$ recursively, and downward in
$m$ and $n$ to compute $Q^+$ recursively, for a total scaling of $\approx \mathcal{O}(J^2N+J(N+M))$.
For each value of $m$, $Q^\pm$ are recursively updated over $n$ until $n_0$ is reached, at which
point the predicted mean can be computed from
\begin{equation}
\bar f^*_i = \sum_{i=1}^{2J} \left[Q^-_{n_0,i}X^-_{m,n_0,i}+Q^+_{n_0+1,i}X^+_{m,n_0+1,i}\right],
\end{equation}
